---
title: "Listing the context in which words appear"
output:
  html_document:
    df_print: paged
author: Christophe@pallier.org
---

Given a text corpus, consisting of a bunch of sentences, a _concordancer_ is a tool that, given a word, displays all the sentences in which it appears.

When the corpus is very large (several millions words), a simple list search is too costly in terms of time.

In this document, I show how to build an efficient concordancer using R. 


## Reading in the Corpus

The corpus is nothing but a text file with one sentence per line. Let us read it in R, using the `scan` function:


```{r}
source('../set-variables.R')
uscorpuspath <- file.path(DATABASES, 'SUBTLEX-US', 'Subtlex-US-corpus.txt.gz')  # nicer if it had been tokenized (?)
Sys.setlocale('LC_ALL','C')


uscorpus <- scan(uscorpuspath, what=character(), sep='\n', encoding='UTF-8', nlines=100000)  
# Notes
#  * here, for demo purposes, we limit ourselves to the first 100k line (about 20% of the entire file)
#  * The fileEncoding should be 'UTF-8', but the subtlex-us file uses another encoding
```


## Building the hashtable

From the corpus, we create an associative array (a.k.a. "hastable") mapping words to lines numbers. The hastable is implemented with an [R environment](http://adv-r.had.co.nz/Environments.html) 

```{r}
assoc <- new.env()
for (linenum in 1:length(uscorpus))
{
  for (token in strsplit(uscorpus[linenum], " ")[[1]])
  { 
    token = gsub('[[:punct:] ]+','', token)  # remove punctuation because the corpus has not been tokenized
    # token = tolower(token)  # To make it case insensitive
    # TODO: to limit the memory footprint. we could use a dictionary to filter out non existing words.
    if (token != "") {
      if (exists(token, assoc)) { 
        assign(token, append(get(token, envir=assoc), linenum), envir=assoc)
      } else {
      assign(token, linenum, envir=assoc)
      }
    }
  }
}

```


## Searching for lines in which a word appears

We can now define a function that returns the lines in which a word appears:

```{r}

get.context <- function(word, envir=assoc) {
   uscorpus[get(word, envir=envir)]
}
```

And use it:

```{r}
print(get.context('dog'))
```


```{r}
save(list = ls(all.names = TRUE), file='uscorp-assoc.RData', envir=assoc)
save(list = c("uscorpus", "get.context"), file='uscorp.RData')
```



